### 2014-01-21 arfi_background.tex
implemented all of Kathy's changes and URLs / citations for ITK-SNAP and 3DSlicer

\subsection{Acoustic Radiation Force Impulse (ARFI) Imaging}
Acoustic radiation force-based elasticity imaging has recently been developed
for investigation of mechanical properties of human soft tissues, including
liver, arteries, heart, prostate, and
breast~\cite{Bouchard2009,Dumont2009,Trahey2004,Dahl2009,proc_congdon04,Palmeri2011,Zhai2010a,hsu07,Palmeri2011a}.
and it provides a unique tool to use to characterize the cervix during
pregnancy.  The acoustic radiation force $\vec{F}$ generated in soft tissues by
focused ultrasound can be described by~\cite{nyborg65,torr84}:
\begin{equation}
\vec{F} = \frac{2 \alpha \vec{I}}{c},
\label{eqn:radforce}
\end{equation}
where $\alpha$ is the acoustic attenuation coefficient of the tissue, $c$ is
the tissue's sound speed, and $\vec{I}$ is the acoustic intensity at a given
point in space.  The acoustic radiation force is generated by a transfer of
momentum from the propagating acoustic wave to the propagation medium through
attenuation mechanisms, such as absorption and scattering of the ultrasonic
wave.  In these studies, the acoustic attenuation and sound speed of the
prostate tissue were assumed to be constant and not vary as a function of the
zonal anatomy, meaning that all displacement differences seen in the prostate
would be assumed to be stiffness-related after normalizing for depth-dependent
focal gain.

>>>>

\subsection{Acoustic Radiation Force Impulse (ARFI) Imaging}
Acoustic radiation force-based elasticity imaging has recently been developed
for investigation of mechanical properties of human soft tissues, including
liver, arteries, heart, prostate, breast, and cervix~\cite{Bouchard2009,Dumont2009,Trahey2004,Dahl2009,proc_congdon04,Palmeri2011,Zhai2010a,hsu07,Palmeri2011a,Palmeri2013}.
The acoustic radiation force $\vec{F}$ generated in soft tissues by
focused ultrasound can be described by~\cite{nyborg65,torr84}:
\begin{equation}
\vec{F} = \frac{2 \alpha \vec{I}}{c},
\label{eqn:radforce}
\end{equation}
where $\alpha$ is the acoustic attenuation coefficient of the tissue, $c$ is
the tissue's sound speed, and $\vec{I}$ is the acoustic intensity at a given
point in space.  The acoustic radiation force is generated by a transfer of
momentum from the propagating acoustic wave to the propagation medium through
attenuation mechanisms, such as absorption and scattering of the ultrasonic
wave.  In these studies, the acoustic attenuation and sound speed of the
prostate tissue were assumed to be constant and not vary as a function of the
zonal anatomy, meaning that all displacement differences seen in the prostate
would be assumed to be stiffness-related after normalizing for depth-dependent
focal gain.

### 2014-01-21 arfi_methods.tex
implemented all of Kathy's changes and URLs / citations for ITK-SNAP and 3DSlicer

For the tracking pulses, 16 parallel receive lines at 5.0 MHz were spaced to
observe both the on and off-axis response of the tissue to the pushing pulses.
Specifically, four lines were dedicated to tracking the on-axis displacement,
with all 4 beams located inside the beamwidth of the pushing pulses such that
the beam spacing was 0.17 mm.

>>>>

For the tracking pulses, 16 parallel receive lines at 5.0 MHz were spaced to
observe both the on and off-axis response of the tissue to the pushing pulses.
Specifically, four lines were dedicated to tracking the on-axis displacement,
with all 4 beams located inside the beamwidth of the pushing pulses such that
the tracking beam spacing was 0.17 mm.

### 2014-01-21 clinical_significance.tex
implemented all of Kathy's changes and URLs / citations for ITK-SNAP and 3DSlicer

PCa screening has led to earlier diagnosis of smaller tumors and more localized
disease; however, it is well known that the sensitivity and specificity of PSA
and DRE are not optimal. In addition, DRE has a low predictive value at lower
PSA ranges, and PSA yields many false
positives.~\cite{Gosselaar2007,Gupta2013,Hricak2007} As such, a theoretical
risk of over-diagnosis and treatment of low-grade, and possibly clinically
insignificant, disease exists.  Moreover, due to the random nature of biopsies,
PCa located outside the routine sampling sites can be missed and the extent
of the cancer might be underestimated.~\cite{Gupta2013,Cornud2012} For example,
in a study by Mufarrij \etal, 45.9--47.2\% of patients who were candidates for
active surveillance, but underwent radical prostatectomy, had a higher Gleason
score on final histopathology than after TRUS biopsy.~\cite{Mufarrij2010} These
inaccuracies may lead to inappropriate diagnosis, imprecise risk assessment and
potentially avoidable morbidity.

>>>>

PCa screening has led to earlier diagnosis of smaller tumors and more localized
disease; however, it is well known that the sensitivity and specificity of PSA
and DRE are not optimal. In addition, DRE has a low predictive value at lower
PSA ranges, and PSA yields many false
positives.~\cite{Gosselaar2007,Gupta2013,Hricak2007} As such, a theoretical
risk of over-diagnosis and treatment of low-grade, and possibly clinically
insignificant, disease exists.  Moreover, due to the random nature of
TRUS-guided systematic biopsies, PCa located outside the routine sampling sites
can be missed and the extent of the cancer might be
underestimated.~\cite{Gupta2013,Cornud2012} For example, in a study by Mufarrij
\etal, 45.9--47.2\% of patients who were candidates for active surveillance,
but underwent radical prostatectomy, had a higher Gleason score on final
histopathology than after TRUS biopsy.~\cite{Mufarrij2010} These inaccuracies
may lead to inappropriate diagnosis, imprecise risk assessment and potentially
avoidable morbidity.

### 2014-01-21 disc_arfi_mr.tex
implemented all of Kathy's changes and URLs / citations for ITK-SNAP and 3DSlicer

ARFI imaging consistently underestimated the extent of the prostate capsule and
central gland apex-to-base dimension relative to MR T2WI
((Figure~\ref{fig:mr_arfi_path_axes}(c), Table~\ref{tab:mr_arfi_axes_error}).
In fact, it is likely the degree of ARFI estimation relative to the actual
apex-to-base prostate length is even greater than suggested by the MR
comparison as this was the dimension with the worst spatial resolution in MR
imaging, with a typical MR T2WI slice thickness of $\sim$3 mm for these
studies.  While higher-resolution MR images could be obtained and analyzed for
more precise comparisons in this imaging dimension, such data were not available
for this study.

>>>>

ARFI imaging consistently underestimated the extent of the prostate capsule and
central gland apex-to-base dimension relative to MR T2WI
((Figure~\ref{fig:mr_arfi_path_axes}(c), Table~\ref{tab:mr_arfi_axes_error}).
In fact, it is likely the degree of ARFI underestimation relative to the actual
apex-to-base prostate length is even greater than suggested by the MR
comparison as this was the dimension with the worst spatial resolution in MR
imaging, with a typical MR T2WI slice thickness of $\sim$3 mm for these
studies.  While higher-resolution MR images could be obtained and analyzed for
more precise comparisons in this imaging dimension, such data were not available
for this study.

### 2014-01-21 fig_mr_arfi_volumes.tex
implemented all of Kathy's changes and URLs / citations for ITK-SNAP and 3DSlicer

\begin{figure}[htb!]
\centering
\begin{tabular}{ccc}
\includegraphics[width=0.3\linewidth]{figs/mr_arfi_volumes} &
\includegraphics[width=0.3\linewidth]{figs/mr_arfi_total_linreg} &
\includegraphics[width=0.3\linewidth]{figs/mr_arfi_central_linreg} \\
(a) CG:PZ Volume & (b) Total Gland Volume & (c) Central Gland Volume \\
\end{tabular}
\caption{Comparison of MR and ARFI zonal anatomy volume estimates from
    manually-segmented images.  Total prostate gland volumes ranged from
    19.6--71.0 cm$^3$ based on MR image models (a), with a moderate correlation
    between MR and ARFI imaging for both the total gland volume (R$^2$ =
    \MRarfiVolTotalRsq, (b)) and the central gland (R$^2$ =
    \MRarfiVolCentralRsq, (c)).  The ARFI total gland volumes had a positive
    bias compared with the MR volumes (\MRarfiVolTotalMeanDiff~$\pm$
    \MRarfiVolTotalStdDiff\% (b)), while the ARFI central zone volumes were on
    average more similar (\MRarfiVolCentralMeanDiff~$\pm$
    \MRarfiVolCentralStdDiff\%, (c)), but with some significant variability
    between study subjects.  Table~\ref{tab:mr_arfi_volumes} contains the
    individual volume estimates for the total prostate and central glands.}
\label{fig:mr_arfi_volumes} 
\end{figure}

>>>>

\begin{figure}[htb!]
\centering
\begin{tabular}{ccc}
\includegraphics[width=0.3\linewidth]{figs/mr_arfi_volumes} &
\includegraphics[width=0.3\linewidth]{figs/mr_arfi_total_linreg} &
\includegraphics[width=0.3\linewidth]{figs/mr_arfi_central_linreg} \\
(a) CG:PZ Volume & (b) Total Gland Volume & (c) Central Gland Volume \\
\end{tabular}
\caption{Comparison of MR and ARFI zonal anatomy volume estimates from
    manually-segmented images.  Total prostate gland volumes ranged from
    19.6--71.0 cm$^3$ based on MR image models (a), with good correlation
    between MR and ultrasound imaging for both the total gland volume (R$^2$ =
    \MRarfiVolTotalRsq, (b)) and the central gland (R$^2$ =
    \MRarfiVolCentralRsq, (c)).  The ultrasound total gland volumes had a positive
    bias compared with the MR volumes (\MRarfiVolTotalMeanDiff~$\pm$
    \MRarfiVolTotalStdDiff\% (b)), while the ARFI central zone volumes were on
    average more similar (\MRarfiVolCentralMeanDiff~$\pm$
    \MRarfiVolCentralStdDiff\%, (c)), but with some significant variability
    between study subjects.  Table~\ref{tab:mr_arfi_volumes} contains the
    individual volume estimates for the total prostate and central glands.}
\label{fig:mr_arfi_volumes} 
\end{figure}

### 2014-01-21 fig_mr_arfi_weight.tex
implemented all of Kathy's changes and URLs / citations for ITK-SNAP and 3DSlicer

\begin{figure}[htb!]
\centering
\begin{tabular}{lll}
\includegraphics[width=0.3\linewidth]{figs/corr_path_vol_weight_vol} &
\includegraphics[width=0.3\linewidth]{figs/corr_weight_vol} &
\includegraphics[width=0.3\linewidth]{figs/corr_pathVol_vol} \\
(a) Path Weight : Path Volume & (b) Image Volume : Prostate Weight & (c) Image Volume : Path Volume \\
\end{tabular}
%\begin{tabular}{ll}
%\includegraphics[width=0.3\linewidth]{figs/corr_weight_vol_no4} &
%\includegraphics[width=0.3\linewidth]{figs/corr_pathVol_vol_no4} \\
%(d) Image Volume : Prostate Weight (-4) & (e) Image Volume : Path Volume (-4) \\
%\end{tabular}
\caption{Tri-axial pathology measurements were used to make an ellipsoidal
    prostate volume approximation based on gross pathology axis measurements,
    which was moderately well-correlated with the excised prostate weights (a,
    R$^2$ = \pathVolWeightRsq).  T2WI MR (blue, X) showed a moderate
    correlation between the reconstructed volumes and prostate weight (R$^2$ =
    \weightMRrsq), while volumes reconstructed from ARFI images (green, O)
    showed weaker correlation (R$^2$ = \weightARFIrsq) (b).  Weaker
    correlations existed between both T2WI MR and ARFI image volumes and
    approximated ellipsoidal prostate pathology volumes (R$^2$ = \pathVolMRrsq
    and \pathVolARFIrsq, respectively) (c).  There was on pathology specimen
    that had an erroneous prostate volume and weight recorded, which has been
    excluded from the linear regressions (magenta circles).}
\label{fig:mr_arfi_weight}
\end{figure}

>>>>

\begin{figure}[htb!]
\centering
\begin{tabular}{lll}
\includegraphics[width=0.3\linewidth]{figs/corr_path_vol_weight_vol} &
\includegraphics[width=0.3\linewidth]{figs/corr_weight_vol} &
\includegraphics[width=0.3\linewidth]{figs/corr_pathVol_vol} \\
(a) Path Weight : Path Volume & (b) Image Volume : Prostate Weight & (c) Image Volume : Path Volume \\
\end{tabular}
%\begin{tabular}{ll}
%\includegraphics[width=0.3\linewidth]{figs/corr_weight_vol_no4} &
%\includegraphics[width=0.3\linewidth]{figs/corr_pathVol_vol_no4} \\
%(d) Image Volume : Prostate Weight (-4) & (e) Image Volume : Path Volume (-4) \\
%\end{tabular}
\caption{Tri-axial pathology measurements were used to make an ellipsoidal
    prostate volume approximation based on gross pathology axis measurements,
    which was moderately well-correlated with the excised prostate weights (a,
    R$^2$ = \pathVolWeightRsq).  T2WI MR (blue, X) showed a moderate
    correlation between the reconstructed volumes and prostate weight (R$^2$ =
    \weightMRrsq), while volumes reconstructed from ARFI images (green, O)
    showed weaker correlation (R$^2$ = \weightARFIrsq) (b).  Weaker
    correlations existed between both T2WI MR and ARFI image volumes and
    approximated ellipsoidal prostate pathology volumes (R$^2$ = \pathVolMRrsq
    and \pathVolARFIrsq, respectively) (c).  There was one pathology specimen
    that had an erroneous prostate volume and weight recorded, which has been
    excluded from the linear regressions (magenta circles).}
\label{fig:mr_arfi_weight}
\end{figure}

### 2014-01-21 future.tex
draft edits and spelling
implemented all of Kathy's changes and URLs / citations for ITK-SNAP and 3DSlicer

\subsection{Future Directions}
Future directions for our research include deeper analysis of our findings,
especially the lateral dimension over-estimation, to improve our anatomic
delineation and assess further its clinical impact.  Technological improvements
to improve anterior prostate boundary delineation in both B-mode and ARFI
imaging are also being pursued and will hopefully allow for the currently
elusive anterior PCa lesions in MR to be seen in ARFI imaging.  3D models of
the prostate and central glands will be used across ARFI and MR imaging
datasets to spatially register them to facilitate correlation of regions of PCa
suspicion between the two modalities, and to correlate with whole-mount
histology of the excised prostate specimens.  This effort will allow the
sensitivity and specificity of each imaging modality for PCa detection and
characterization to be quantified.  While ARFI imaging of the prostate in a
constantly evolving technology, the fact that one can view the different
prostatic zones on ARFI images in addition to MRI is encouraging and could lead
to targeted diagnostic biopsies and therapies with real-time imaging for men
with PCa.

>>>>

\subsection{Future Directions}
Future directions for our research include deeper analysis of our findings,
especially the lateral dimension over-estimation, to improve our anatomic
delineation and assess further its clinical impact.  Technological improvements
to improve anterior prostate boundary delineation in both B-mode and ARFI
imaging are also being pursued for better visualization in this region.  3D
models of the prostate and central glands will be used across ARFI and MR
imaging datasets to spatially register them to facilitate correlation of
regions of PCa suspicion between the two modalities, and to correlate with
whole-mount histology of the excised prostate specimens.  This effort will
allow the sensitivity and specificity of each imaging modality for PCa
detection and characterization to be quantified.  Additionally, we will be
pursuing quantitative measurement of the pressure being applied during imaging
to help reduce prostate deformation and reduce the impact of prostate stiffness
nonlinearity in our displacement images.  While ARFI imaging of the prostate is
a constantly evolving technology, the fact that one can view the different
prostatic zones on ARFI images in addition to MRI is encouraging and could lead
to targeted diagnostic biopsies and therapies with real-time imaging for men
with PCa.

### 2014-01-21 image_proc_methods.tex
implemented all of Kathy's changes and URLs / citations for ITK-SNAP and 3DSlicer

\subsection{Image Zonal Anatomy Segmentation and 3D Model Rendering}
\subsubsection{MR Image Segmentation and Modeling}
Axial MR T2WI images were manually segmented using the smooth polygon tool in
ITK-SNAP~\cite{Yushkevich2006}, using unique labels for the PZ, CG and AFS. The
gland was segmented from base-to-apex.  The base was identified below the
bladder, and subsequent images were segmented until the last slice with visible
prostatic tissue was identified caudally. The CG, urethra, PZ and AFS were
segmented independently according to their well-established anatomical
characteristics on
T2WI.~\cite{Barentsz2012,Jung2012,Poon1985,Hricak2007,Bonekamp2011} The PZ was
identified by its homogeneous high signal intensity on T2WI, which is usually
similar to that of the nearby peri-prostatic fat. The CG was visualized and
delineated based on its heterogeneous, lower signal intensity and its location
(Figure~\ref{fig:mr_anatomy}). The urethra was included in the CG segmentation.
When visible, the AFS was identified by its low T2 signal intensity and its
location anterior to the CG.

@@

Segmented image stacks were imported into 3D Slicer (v4.3.0) and 3D
models were rendered using the following parameters (Table~\ref{tab:3dslicer}):

>>>>

\subsection{Image Zonal Anatomy Segmentation and 3D Model Rendering}
\subsubsection{MR Image Segmentation and Modeling}
Axial MR T2WI images were manually segmented using the smooth polygon tool in
ITK-SNAP (\url{http://www.itksnap.org})~\cite{Yushkevich2006}, using unique labels for the PZ, CG and AFS. The
gland was segmented from base-to-apex.  The base was identified below the
bladder, and subsequent images were segmented until the last slice with visible
prostatic tissue was identified caudally. The CG, urethra, PZ and AFS were
segmented independently according to their well-established anatomical
characteristics on
T2WI.~\cite{Barentsz2012,Jung2012,Poon1985,Hricak2007,Bonekamp2011} The PZ was
identified by its homogeneous high signal intensity on T2WI, which is usually
similar to that of the nearby peri-prostatic fat. The CG was visualized and
delineated based on its heterogeneous, lower signal intensity and its location
(Figure~\ref{fig:mr_anatomy}). The urethra was included in the CG segmentation.
When visible, the AFS was identified by its low T2 signal intensity and its
location anterior to the CG.

@@

Segmented image stacks were imported into 3D Slicer (\url{http://www.slicer.org}, v4.3.0)~\cite{Fedorov2012} and 3D
models were rendered using the following parameters (Table~\ref{tab:3dslicer}):

### 2014-01-21 mr_background.tex
removed unnecessary higher-order MR background
implemented all of Kathy's changes and URLs / citations for ITK-SNAP and 3DSlicer

\subsection{MR Imaging in Prostate Diagnostics}
MR imaging has been available for use in the workup of patients with PCa since
the early 1980s, but the early work on its diagnostic accuracy is heterogeneous
due to heavy reliance on morphology via T1 and T2-weighted imaging (T2WI). More
recent abilities to include not only anatomic, but also biologic and functional
dynamic parameters, into MR analysis via diffusion-weighted imaging (DWI),
dynamic contrast-enhanced (DCE) imaging or MR spectroscopic imaging (MRSI) is
promising in the future diagnosis and management of PCa.

@@

The prostate can be divided into glandular and non-glandular components. The
glandular components include the peripheral zone (PZ) and the central gland,
which are typically easily distinguishable on T2WI. The central gland includes
the central zone, transition zone and the periurethral glandular
tissue.~\cite{Jung2012} Other anatomical markers such as the urethra,
verumontanum and ejaculatory are also often seen on T2WI.

@@

%DWI is based on the free movement of water particles in tissue and measures the
%degree of motion restriction.~\cite{Koh2007} Normal prostatic tissue is very
%glandular with plenty of water molecule movement. On the other hand, tumors
%have high cellular density and restricted water movement, which leads to
%decreased diffusion. Due to this restricted diffusion, tumors seem to be of
%higher intensity on DWI.~\cite{Koh2007}
%
%DCE-MRI provides a measure of tumor vascularity and takes advantage of
%tumor-induced angiogenesis in PCa.~\cite{Noworolski2005} Increased vascularity
%in PCa results in earlier and higher peak contrast enhancement and a faster
%washout when compared to normal prostatic tissue. These known characteristics
%of PCa are helpful when trying to localize lesions. 
%
%Like DWI, MRSI can be helpful in lesion characterization as it provides
%information on the presence of certain metabolites in
%tissue.~\cite{Barentsz2012} Choline and citrate are especially useful in the
%setting of PCa. Choline is critical in cell membrane synthesis and is thus
%usually elevated in cells with cancerous behavior.~\cite{Gupta2013} Healthy
%prostate epithelium synthesizes and secretes large quantities of citrate but
%levels are decreased in PCa.~\cite{Costello1999,Kurhanewicz2002} Therefore, an
%increase in the choline-to-citrate ratio on MRSI can be used as an indicator of
%malignancy.~\cite{Bonekamp2011}

>>>>

\subsection{MR Imaging in Prostate Diagnostics}
MR imaging has been available for use in the workup of patients with PCa since
the early 1980s, but the early work on its diagnostic accuracy is heterogeneous
due to heavy reliance on morphology via T1 and T2-weighted imaging (T2WI). More
recent abilities to include not only anatomic, but also biologic and functional
dynamic parameters, into MR analysis via diffusion-weighted imaging (DWI),
dynamic contrast-enhanced (DCE) imaging or MR spectroscopic imaging (MRSI) are
promising in the future diagnosis and management of PCa.

@@

The prostate can be divided into glandular and non-glandular components. The
glandular components include the peripheral zone (PZ) and the central gland,
which are typically easily distinguishable on T2WI.  Other anatomical markers
such as the urethra, verumontanum and ejaculatory are also often seen on T2WI.

@@

### 2014-01-21 results.tex
implemented all of Kathy's changes and URLs / citations for ITK-SNAP and 3DSlicer

\section{Results}\label{sect:results}
The subvolumes associated with the zonal anatomy in each imaging modality were
measured (Figure~\ref{fig:mr_arfi_volumes}(a)), with moderate correlation
between the ARFI and MR total prostate gland volumes (R$^2$ =
\MRarfiVolTotalRsq), with a mean overestimation of of
\MRarfiVolTotalMeanDiff~$\pm$ \MRarfiVolTotalStdDiff\% by ARFI imaging compared
to MR volumes (Figure~\ref{fig:mr_arfi_volumes}(b)).  Central gland volumes
were well-correlated between ARFI and MR images (R$^2$ = \MRarfiVolCentralRsq)
with no significant mean over/under-estimation (\MRarfiVolCentralMeanDiff~$\pm$
\MRarfiVolCentralStdDiff\%), but significant variability between the cases
(Figure~\ref{fig:mr_arfi_volumes}(c)).  Table~\ref{tab:mr_arfi_volumes} has the
individually-measured volumes for MR and ARFI imaging for each study subject.

>>>>

\section{Results}\label{sect:results}
The subvolumes associated with the zonal anatomy in each imaging modality were
measured (Figure~\ref{fig:mr_arfi_volumes}(a)), with good correlation
between the ultrasound and MR total prostate gland volumes (R$^2$ =
\MRarfiVolTotalRsq), with a mean overestimation of of
\MRarfiVolTotalMeanDiff~$\pm$ \MRarfiVolTotalStdDiff\% by ultrasound imaging compared
to MR volumes (Figure~\ref{fig:mr_arfi_volumes}(b)).  Central gland volumes
were well-correlated between ARFI and MR images (R$^2$ = \MRarfiVolCentralRsq)
with no significant mean over/under-estimation (\MRarfiVolCentralMeanDiff~$\pm$
\MRarfiVolCentralStdDiff\%), but significant variability between the cases
(Figure~\ref{fig:mr_arfi_volumes}(c)).  Table~\ref{tab:mr_arfi_volumes} has the
individually-measured volumes for MR and ARFI imaging for each study subject.
